Greenhouse gas emissions model

### 2021-07-13 210617_GHG_model.Rmd
Add references and format coed with notes

---
title: "GHG model"
author: "Cory and Eike"
date: "6/17/2021"
output: html_document
---

@@

## Retracing the equations of the GHG emissions model

We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport, and to possibly optimize the model.

Here's the original model

@@

library(decisionSupport)
#CW Note ####
# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
# This project will look in relative file paths (i.e. same or lower folders)
input_table <- "livestock_ghg_input_table.csv"
results_folder <- "MC results"

@@

# Enteric fermentation emissions ($CH_4$)
  #CW Note ####
  #in these code sections I am missing the overview. 
  # it is not immediately clear what this is meant to do... 
  # from a specific publication?

# This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
# We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.

@@

# here, C is an animal type coefficient
# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)
# SWG is the "Shrunk Weight Gain" in kg per day - we're not sure what exactly this means.

#This used to be the following equation, which we didn't find a source for:
#NE_growth<-22.02*(((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*(weight_gain_WG^1.097))

#The code doesn't include net energy mobilization by animals that lose weight (p. 314)

@@

#in Joshua's equation the 0.10 was a variable called pregnancy_coefficient_Cp with a value around 0.03. We're not sure where this came from.

@@

#Equation 11 contains calculation of the metabolizable energy, but that doesn't seem to be used anywhere

@@

#the following equation usually contains the energy mobilized by animals losing weight. This wasn't calculated, so we set this to 0
NE_mobilized<-0
#same for extra energy spent by working animals
NE_w<-0

@@

#Y_m here is the methane conversion rate - Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)

# Compute the CO2 equation for enteric methane
#This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CO2=enteric_CH4*25

@@

# The maximum methane emission from manure (Bo) depends on the animal type. In the input table, we
# treat these as constants for each animal type. The following line assigns the corresponding Bo value
# to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
# animal types 4-6 and 0.135 for animal types 1-3 and 7-12.

@@

# Estimate CO2eq for manure methane
### this one also contained a multiplication by 365 that seemed erroneous
mm_CO2<-mm_ch4*25

@@

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

---
title: "Greenhouse gas emissions model"
author: "Cory and Eike"
bibliography:
 - references/packages.bib
 - references/papers.bib
output: html_document
---

@@

```{r, warning = F, include = F}
#install and load packages
library(decisionSupport)
library(rmarkdown)

#Automatically write R package citation entries to a .bib file
knitr::write_bib(c(.packages(), 
                   'decisionSupport'), 'references/packages.bib')

## Generating the GHG emissions model

We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport [@R-decisionSupport], and to possibly optimize the model.

The formulas come from Chapter 10 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories [@dong_ipcc_2006]. The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).  Many variables are based on the work done in 'Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya' [@wilkes_variation_2020] and in 'Methods and guidance to support MRV of livestock emissions' [@wilkes_methods_2019].

*Here are our changes and additions to the original model:*

@@

library(decisionSupport)
###CW Note ####
# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
###CW Note ####
### # This project will be stored in relative file paths (i.e. same or lower folders)
input_table <- "livestock_ghg_input_table.csv"

@@

# Enteric fermentation emissions ($CH_4$)
  ####CW Note ####
  #in these code sections I am missing the overview. 
  # it is not immediately clear what this is meant to do... 
  # from a specific publication?

# This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
#####EL CW Note ####
## We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.

@@

# here, C is an animal type coefficient
# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)

##### EL CW Note ####
# SWG is the "Shrunk Weight Gain" in kg per day - we're not sure what exactly this means.

##### EL CW Note #### 
#This used to be the following equation, which we didn't find a source for:
#NE_growth<-22.02*(((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*(weight_gain_WG^1.097))

##### EL CW Note #### 
#The code doesn't include net energy mobilization by animals that lose weight (p. 314)

@@

##### EL CW Note #### 
#in Joshua's equation the 0.10 was a variable called pregnancy_coefficient_Cp with a value around 0.03. We're not sure where this came from.

@@

##### EL CW Note #### 
#Equation 11 contains calculation of the metabolizable energy, but that doesn't seem to be used anywhere

@@

##### EL CW Note #### 
#the following equation usually contains the energy mobilized by animals losing weight. This wasn't calculated, so we set this to 0
NE_mobilized<-0
#same for extra energy spent by working animals
NE_w<-0

@@

##### EL CW Note #### 
#Y_m here is the methane conversion rate - Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)

# Compute the CO2 equation for enteric methane
##### EL CW Note #### 
##### #This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CO2=enteric_CH4*25

@@

# The maximum methane emission from manure (Bo) depends on the animal type. In the input table, we
# treat these as constants for each animal type. The following line assigns the corresponding Bo value
##### EL CW Note #### 
##### # to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
# animal types 4-6 and 0.135 for animal types 1-3 and 7-12.

@@

# Estimate CO2eq for manure methane
### ##### EL CW Note #### 
### this one also contained a multiplication by 365 that seemed erroneous
mm_CO2<-mm_ch4*25

@@

This document was generated using R Markdown [@R-rmarkdown].

# References
